This section defines the general operation of a single integer lifting
operation.

Each lifting process applies either to odd or to even coefficients. If
it applies to even coefficients, then the even coefficients are modified
using the values of odd coefficients only. If it applies to odd
coefficients, then the odd coefficients are modified using the values of
even coefficients only. Specifically, a single lifting filtering
operation is of the form:

\begin{displaymath}
  C(2n) = C(2n) + \big( \sum^M_{i=-N} t_i C(2(n+i) - 1) \big) >> s
\end{displaymath}

if it operates on even coefficients and of the form

\begin{displaymath}
  C(2n+1) = C(2n+1) + \big( \sum^M_{i=-N} t_i C(2(n+i)) \big) >> s
\end{displaymath}

if it operates on odd coefficients. The values ti are the lifting filter
tap values; the value s is the lifting filter scale factor.

The lifting operation is applied for all the even, or all the odd,
coefficients in the row or column array C. A decoder may perform the
individual coefficient filtering operations in any order, as this does
not affect the result.

The filtering process uses edge extension.  Where filtering requires
values which fall out of the range of values of the array C, then the
value selected is determined as follows:

- C[2k] is identified with C[0] if $2k < 0$
- C[2k] is identified with C[length(C)-2] if $2k > length(C)$
- C[2k+1] is identified with C[1] if $2k+1 < 0$
- C[2k+1] is identified with C[length(C)-1] if $2k + 1 > length(C)$

\begin{informative}
This specification defines the lifting process on the basis of lifting
procedures applied to an entire row or column consecutively. It is
possible to implement lifting filtering operations so that a filtering
operation associated with one lifting filter is followed by a filtering
operation associated with another lifting filter. I.e. the order of
iteration is changed. In this case, the order in which filtering is
applied to coefficients does affect the outcome of the process as even
lifting operations may be followed by odd ones, and care must be taken
that values are not modified in the wrong order. Nevertheless such an
implementation may be more efficient, and complies with this
specification if it produces identical results.
\end{informative}

